SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-10-26 maintex.tex
replaced Section with \S

% Organization
This paper is organized as follows; in Section \ref{sec:data} we will discuss the properties of the MaNGA survey along with the construction of our pair and control samples, in Section \ref{sec:analysis} we will discuss how we measure star formation rates and how we build radial profiles of star formation, in Section \ref{sec:results} we study the radial profiles as a function of stellar mass, projected separation, and the mass ratio, in Section \ref{sec:disc} we compare our work against previous works, and in Section \ref{sec:sum} we summarize the findings of the work. 
% cosmology
Throughout we adopt the $\Lambda$CDM cosmology with $\Omega_{\rm m}=0.3$, $\Omega_\Lambda=0.7$, and $h=0.7$.

@@

Among the 6142 MaNGA IFUs considered here, 404 have multiple galaxy objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta v$ $<$ 500 km s$^{-1}$ to remove projected companions. Given the redshift range of the MaNGA sample and the size of MaNGA's IFUs, the maximum projected separation for a companion galaxy in the IFU is $\sim$40 kpc. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming as described in Section \ref{sec:data}. These requirements reduce our sample down to 87 star forming MaNGA targets with inside-IFU companions. While the companions to the inside-IFU MaNGA targets contained are within the MaNGA IFUs, we only use the MaNGA targets when building radial profiles. We leave out the identified companions from the analysis because the MaNGA targets are covered by the NSA catalog with geometries and total stellar masses while the companions are not covered by the NSA catalog.

@@

In the first method, we simply control for the stellar mass between the pairs and controls. We split both the pair and control samples into five evenly spaced stellar mass bins over the range, \logm\ $=$ 9.0$-$11.5. The pairs are then compared to their respective controls within each stellar mass bins. While the redshift is not constrained in this method, we will show in Section \ref{sec:mass-bin} that this method is sufficient to reveal the merger induced star formation.

@@

With the individual log(sSFR) profiles built for each MaNGA galaxy, we now use two different methods to compare the profiles of paired galaxies against the profiles of control galaxies. In the first method, in Section \ref{sec:mass-bin}, paired galaxies are compared to control galaxies within evenly spaced stellar mass-bins. In the second method, in Section \ref{sec:tailored}, paired galaxies are compared to a subset of 20 control galaxies which are matched in stellar mass and redshift.

@@

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_dm.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by the mass ratio of the pair. The inside-IFU sample is left out of this analysis because we do not yet have reliable measurements of their total stellar mass ratios.}
\label{fig:ssfr_dm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the second method, we match each paired galaxy to a set of 20 control galaxies of similar stellar masses and redshifts, as described in Section \ref{sec:control}. We then take the median of the azimuthally averaged profiles of the tailored control sample. Finally, we obtain the $\Delta\log$sSFR profile for each of the 175 paired galaxies by calculating the difference between its sSFR profile and the median profile of its control sample.

@@

The difference profiles are shown to be centrally enhanced by $\sim$0.25 $\pm$ 0.1 dex. Similar to what was seen in Section \ref{sec:mass-bin}, the level of the enhancement is independent of galaxy mass with the exception of the highest mass bin, \logm\ $=$ 11.0$-$11.5, which is 0.15$-$0.25 dex higher than the sample average.

The central enhancement to the sSFR falls to zero around 1.2 \reff\ and the outskirts of the galaxies are lightly suppressed by 0.1 dex. The paired galaxies in the mass range, \logm\ $=$ 9.0$-$9.5, feature enhanced sSFR in their disks of 0.1$-$0.2 dex. As we will show in the next section, Section \ref{sec:dm}, the offsets seen in the disks of paired galaxies are largely driven by whether the galaxy is the more massive or less massive companion in the pair.

To understand why these high mass galaxies behave differently from the rest of the sample, we compare the distribution of the projected separation for the whole sample and for galaxies in the highest mass bin in Figure \ref{fig:sep_hist}. The whole sample has a v-shaped separation distribution. This is the result of the combination of the two different samples, the inside-IFU and outside-IFU samples. The outside-IFU sample should find more pairs at wide projected separations due to the larger volume to find companions in. The separation distribution of the inside-IFU sample is limited by MaNGA's IFU sizes and the sample's redshift distribution; the maximum possible projected separation for a MaNGA target with $z$ $=$ 0.15 and a 127 spaxel IFU is about 40 kpc. The highest mass galaxies in the sample are at closer projected separations in comparison to the rest of the sample with 70\% of the highest mass galaxies being within 20 kpc. As we will show in Section \ref{sec:sep}, the central sSFR enhancement is strongest at small projected separation which means that higher sSFR in the centers of the massive galaxies may be driven by their close projected separations.

The results we obtain from the tailored-control method is largely consistent with the mass-binning method. Between the two methods we see that the $\Delta$log(sSFR) profile is independent of the total stellar mass of the galaxies. In the next two sections, we will use the tailored-control method to study how the difference profiles behave as a function of the mass ratio (Section \ref{sec:dm}) and projected separation (Section \ref{sec:sep}).

>>>>

% Organization
This paper is organized as follows; in \S~\ref{sec:data} we will discuss the properties of the MaNGA survey along with the construction of our pair and control samples, in \S~\ref{sec:analysis} we will discuss how we measure star formation rates and how we build radial profiles of star formation, in \S~\ref{sec:results} we study the radial profiles as a function of stellar mass, projected separation, and the mass ratio, in \S~\ref{sec:disc} we compare our work against previous works, and in \S~\ref{sec:sum} we summarize the findings of the work. 
% cosmology
Throughout we adopt the $\Lambda$CDM cosmology with $\Omega_{\rm m}=0.3$, $\Omega_\Lambda=0.7$, and $h=0.7$.

@@

Among the 6142 MaNGA IFUs considered here, 404 have multiple galaxy objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta v$ $<$ 500 km s$^{-1}$ to remove projected companions. Given the redshift range of the MaNGA sample and the size of MaNGA's IFUs, the maximum projected separation for a companion galaxy in the IFU is $\sim$40 kpc. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming as described in \S~\ref{sec:data}. These requirements reduce our sample down to 87 star forming MaNGA targets with inside-IFU companions. While the companions to the inside-IFU MaNGA targets contained are within the MaNGA IFUs, we only use the MaNGA targets when building radial profiles. We leave out the identified companions from the analysis because the MaNGA targets are covered by the NSA catalog with geometries and total stellar masses while the companions are not covered by the NSA catalog.

@@

In the first method, we simply control for the stellar mass between the pairs and controls. We split both the pair and control samples into five evenly spaced stellar mass bins over the range, \logm\ $=$ 9.0$-$11.5. The pairs are then compared to their respective controls within each stellar mass bins. While the redshift is not constrained in this method, we will show in \S~\ref{sec:mass-bin} that this method is sufficient to reveal the merger induced star formation.

@@

With the individual log(sSFR) profiles built for each MaNGA galaxy, we now use two different methods to compare the profiles of paired galaxies against the profiles of control galaxies. In the first method, in \S~\ref{sec:mass-bin}, paired galaxies are compared to control galaxies within evenly spaced stellar mass-bins. In the second method, in \S~\ref{sec:tailored}, paired galaxies are compared to a subset of 20 control galaxies which are matched in stellar mass and redshift.

@@

\subsubsection{Tailored-Controls Difference Profiles}\label{sec:tailored}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_dm.pdf}
\caption[]{Same as Figure \ref{fig:ssfr_mass} except the difference profiles are split by the mass ratio of the pair. The inside-IFU sample is left out of this analysis because we do not yet have reliable measurements of their total stellar mass ratios.}
\label{fig:ssfr_dm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the second method, we match each paired galaxy to a set of 20 control galaxies of similar stellar masses and redshifts, as described in \S~\ref{sec:control}. We then take the median of the azimuthally averaged profiles of the tailored control sample. Finally, we obtain the $\Delta\log$sSFR profile for each of the 175 paired galaxies by calculating the difference between its sSFR profile and the median profile of its control sample.

@@

The difference profiles are shown to be centrally enhanced by $\sim$0.25 $\pm$ 0.1 dex. Similar to what was seen in \S~\ref{sec:mass-bin}, the level of the enhancement is independent of galaxy mass with the exception of the highest mass bin, \logm\ $=$ 11.0$-$11.5, which is 0.15$-$0.25 dex higher than the sample average.

The central enhancement to the sSFR falls to zero around 1.2 \reff\ and the outskirts of the galaxies are lightly suppressed by 0.1 dex. The paired galaxies in the mass range, \logm\ $=$ 9.0$-$9.5, feature enhanced sSFR in their disks of 0.1$-$0.2 dex. As we will show in the next section, \S~\ref{sec:dm}, the offsets seen in the disks of paired galaxies are largely driven by whether the galaxy is the more massive or less massive companion in the pair.

To understand why these high mass galaxies behave differently from the rest of the sample, we compare the distribution of the projected separation for the whole sample and for galaxies in the highest mass bin in Figure \ref{fig:sep_hist}. The whole sample has a v-shaped separation distribution. This is the result of the combination of the two different samples, the inside-IFU and outside-IFU samples. The outside-IFU sample should find more pairs at wide projected separations due to the larger volume to find companions in. The separation distribution of the inside-IFU sample is limited by MaNGA's IFU sizes and the sample's redshift distribution; the maximum possible projected separation for a MaNGA target with $z$ $=$ 0.15 and a 127 spaxel IFU is about 40 kpc. The highest mass galaxies in the sample are at closer projected separations in comparison to the rest of the sample with 70\% of the highest mass galaxies being within 20 kpc. As we will show in \S~\ref{sec:sep}, the central sSFR enhancement is strongest at small projected separation which means that higher sSFR in the centers of the massive galaxies may be driven by their close projected separations.

The results we obtain from the tailored-control method is largely consistent with the mass-binning method. Between the two methods we see that the $\Delta$log(sSFR) profile is independent of the total stellar mass of the galaxies. In the next two sections, we will use the tailored-control method to study how the difference profiles behave as a function of the mass ratio (\S~\ref{sec:dm}) and projected separation (\S~\ref{sec:sep}).
